---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(tidyverse)
#library(CelltypeR)
```



```{r}
seu <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/Parse48_analysisMaria.RDS")

unique(seu$sample)

Idents(seu) <- "sample"
levels(seu)

samples <- levels(seu)
samples

```

Add meta data information

```{r}
# using celltypeR function

# add the culture type NPC, DA neurons 2weeks, Motor neurons
new_vector <- ifelse(grepl("NPC", samples), "NPC",
                     ifelse(grepl("Final", samples), "Neurons2weeks",
                            ifelse(grepl("^M", samples), "MotorNeurons", "Other")))

# Print the new vector
print(new_vector)

seu <- annotate(seu, annotations = new_vector, to_label = "sample",annotation_name = "CultureType")

DimPlot(seu)

# check with sample names

table(seu$sample, seu$CultureType)


```

Add the lines and then the diseases status

```{r}

  # Create a new vector with line names
new_line_vector <- gsub("^(NPC|Final)", "", samples)
new_line_vector <- gsub("3123$", "3123-iso", new_line_vector)
new_line_vector[new_line_vector == "P65P"] <- "3123-isoP65P"
new_line_vector[new_line_vector == "Q65Q"] <- "3123-isoQ65Q"

# Print the new vector with line names


new_line_vector[new_line_vector == "2AIW002-02"] <- "AIW002-02"

print(new_line_vector)

seu <- annotate(seu, annotations = new_line_vector, to_label = "sample",annotation_name = "Line")


# check
table(seu$sample, seu$Line)

```
```{r, fig.width=6}
DimPlot(seu)
```


```{r}
# add disease status

# Create the lookup table for Line to DiseaseStatus mapping
line_to_disease <- c("3448" = "HC",
                     "TD22" = "HC",
                     "AIW001-02" = "HC",
                     "AIW002-02" = "HC",
                     "2965" = "PD",
                     "3383" = "PD",
                     "3575" = "PD",
                     "TD06" = "PD",
                     "TD07" = "PD",
                     "3123" = "PD",
                     "3567" = "PD",
                     "3595" = "PD",
                     "3940" = "PD",
                     "3567-iso" = "PD-iso",
                     "3940R-iso" = "PD-iso",
                     "TD06-iso" = "PD-iso",
                     "3940N-iso" = "PD-iso",
                     "3123-isoQ65Q" = "PD-iso",
                     "3123-isoP65P" = "PD-iso",
                     "3123-iso" = "PD-iso",
                     "3383-iso" = "PD-iso",
                     "3575-iso" = "PD-iso",
                     "3595-iso" = "PD-iso")

# Assuming "Line" is your starting character vector
Idents(seu) <- "Line"
Line <- levels(seu)
# Create a new vector of disease statuses using the lookup table
disease_status_vector <- ifelse(Line %in% names(line_to_disease), 
                                line_to_disease[Line], 
                                "other")

# Print the new vector of disease statuses
print(disease_status_vector)

seu <- annotate(seu, annotations = disease_status_vector, to_label = "Line",annotation_name = "DiseaseStatus")


# check
table(seu$Line, seu$DiseaseStatus)


```

```{r}
cellcounts <- as.data.frame(table(seu$Line,seu$CultureType))
table(seu$Line,seu$CultureType,seu$DiseaseStatus)

# Convert long_df to the wide format
wide_df <- pivot_wider(cellcounts, 
                       id_cols = Var1, 
                       names_from = Var2, 
                       values_from = Freq)

write.csv(wide_df,"cellcounts_sample.csv")

```

```{r}
DimPlot(seu, group.by = "DiseaseStatus")

```




Subset for to remove pools and Maria's samples

```{r}
unique(seu$sample)
Idents(seu) <- "DiseaseStatus"

seu.n <- subset(seu, idents = c("HC","PD","PD-iso"))
dim(seu.n)
dim(seu)

DimPlot(seu.n)

```
Reprocess subset

```{r}

seu.n <- RunPCA(seu.n, npcs = 20, verbose = FALSE)
seu.n <- RunUMAP(seu.n, reduction = "pca", dims = 1:20, n.neighbors = 81)
DimPlot(seu.n)
seu.n$CultureType
DimPlot(seu.n, group.by = "CultureType")
DimPlot(seu.n, group.by = "Line")

```

```{r}
seu.n <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamples.RDS")
  
```



Get clusters

```{r}
seu.n <- NormalizeData(seu.n)
seu.n <- FindVariableFeatures(seu.n)
seu.n <- FindNeighbors(seu.n, dims = 1:20, k.param = 81)
seu.n <- FindClusters(seu.n, resolution = 0.3)
DimPlot(seu.n, group.by = "RNA_snn_res.0.3")
#seu.n$RNA_snn_res.0.3
DimPlot(seu.n, group.by = "CultureType")

```





Subset out neurons and NPC

```{r}

#seu.n$CultureType
Idents(seu.n) <- "CultureType"
neurons <- subset(seu.n, idents = "Neurons2weeks")

Idents(seu.n) <- "CultureType"
NPC <- subset(seu.n, idents = "NPC")

```

```{r}

seu <- neurons
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:20, n.neighbors = 81)
seu <- FindVariableFeatures(seu)
seu <- FindNeighbors(seu, dims = 1:20, k.param = 81)
seu <- FindClusters(seu, resolution = 0.3)
DimPlot(seu)
DimPlot(seu, group.by = "CultureType")
DimPlot(seu, group.by = "Line")

neurons <- seu

```



```{r}
seu <- NPC
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:20, n.neighbors = 81)
seu <- FindVariableFeatures(seu)
seu <- FindNeighbors(seu, dims = 1:20, k.param = 81)
seu <- FindClusters(seu, resolution = 0.3)
DimPlot(seu)
DimPlot(seu, group.by = "CultureType")
DimPlot(seu, group.by = "Line")

NPC <- seu

DimPlot(NPC, group.by = "DiseaseStatus")

```



save objects 
```{r}

saveRDS(seu, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/Parse48_allSamples.RDS")

saveRDS(seu.n, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamples.RDS")
saveRDS(neurons, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeurons.RDS")
saveRDS(NPC, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNPC.RDS")


neurons <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeurons.RDS")


```

```{r}
table(neurons$Line)
Idents(neurons) <- "Line"
levels(neurons)
table(neurons$DiseaseStatus,neurons$Line)

```

Now integrate samples

```{r}

# All samples is too large to integrate 

Idents(neurons) <- "DiseaseStatus"
levels(neurons)
neur.PD <- subset(neurons, idents = "PD")
neur.PDiso <- subset(neurons, idents = "PD-iso")
neur.HC <- subset(neurons, idents = "HC")

# make a list of seurat objects by our cell type variable
# will integrate the PD, PD-iso, HC separately and then merge them will see if will integrate or not

```

```{r}
# an integrate function

# Define a function for integrating a list of Seurat objects
integrate_seurat_objects <- function(seurat_list, dims = 1:30) {
  # Normalize and find variable features for each object
  for (i in 1:length(seurat_list)) {
    seurat_list[[i]] <- NormalizeData(seurat_list[[i]], verbose = FALSE)
    seurat_list[[i]] <- FindVariableFeatures(seurat_list[[i]], selection.method = "vst")
  }

  # Create an empty Seurat object to store the integrated data
  integrated_seurat <- subset(seurat_list[[1]])

  # Iterate over the list of Seurat objects
  for (i in 1:length(seurat_list)) {
    # Rename the 'orig.ident' metadata inside the Seurat object to match the object name in the list
    seurat_list[[i]]$orig.ident <- names(seurat_list)[i]
  }

  sample.list <- seurat_list
  for (i in 1:length(sample.list)) {
    # Normalize and scale the data
    sample.list[[i]] <- NormalizeData(sample.list[[i]], verbose = FALSE)
    sample.list[[i]] <- ScaleData(sample.list[[i]], verbose = FALSE)
    # Find variable features
    sample.list[[i]] <- FindVariableFeatures(sample.list[[i]], selection.method = "vst")
    # Get the variable features
    variable_features <- VariableFeatures(sample.list[[i]])
    # Run PCA with the variable features
    sample.list[[i]] <- RunPCA(sample.list[[i]], verbose = FALSE, npcs = 30, features = variable_features)
  }

  # Find integration anchors
  int.anchors <- FindIntegrationAnchors(object.list = sample.list, dims = dims, reduction = "rpca")

  # Integrate data
  integrated_seurat <- IntegrateData(anchorset = int.anchors, dims = dims)

  return(integrated_seurat)
}



```

Use function to integrate the HC

```{r}

sublist <- SplitObject(neur.HC, split.by = "Line")

int.HC <- integrate_seurat_objects(sublist, dims = 1:30)
rm(sublist)
sublist <- SplitObject(neur.PD, split.by = "Line")

int.PD <- integrate_seurat_objects(sublist, dims = 1:30)
rm(sublist)
sublist <- SplitObject(neur.PDiso, split.by = "Line")

int.PDiso <- integrate_seurat_objects(sublist, dims = 1:30)


```



```{r}

# Create a list of the three Seurat objects
DefaultAssay(int.HC) <- "RNA"
DefaultAssay(int.PD) <- "RNA"
#DefaultAssay(int.PDiso) <- "RNA" # didn't save
Neuron_list <- list("HC" = int.HC,"PD" = int.PD)


# Integrate the Seurat objects
integrated_seurat <- integrate_seurat_objects(Neuron_list, dims = 1:20)



```

```{r}
# save integrated objects
saveRDS(int.HC, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintHC.RDS")
saveRDS(int.PD, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintPD.RDS")
saveRDS(int.PDiso, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintPDiso.RDS")

```

Read in the integrated objects and then integrate all 3

```{r}

int.HC <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintHC.RDS")
int.PD <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintPD.RDS")
int.PDiso("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeuronsintPD.RDS")

```

Integrated only the PD and HC


```{r}
# make UMAP from integrated
seu <- integrated_seurat
seu <- ScaleData(seu)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:20, n.neighbors = 81)
seu <- FindVariableFeatures(seu)
seu <- FindNeighbors(seu, dims = 1:20, k.param = 81)
seu <- FindClusters(seu, resolution = 0.3)
DimPlot(seu)
DimPlot(seu, group.by = "Line")
DimPlot(seu, group.by = "DiseaseStatus")


```


Samples are not well integrated.  I'll try to integrate by lines

```{r}

# subset just disease and control
Idents(neurons) <- "DiseaseStatus"
Neu.HCPD <- subset(neurons, idents = c("HC","PD"))
sublist <- SplitObject(Neu.HCPD, split.by = "Line")



# Integrate the Seurat objects
int.PDHC <- integrate_seurat_objects(sublist, dims = 1:20)



```

```{r}
# make UMAP from integrated
seu <- int.PDHC
seu <- ScaleData(seu)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:20, n.neighbors = 81)
seu <- FindVariableFeatures(seu)
seu <- FindNeighbors(seu, dims = 1:20, k.param = 81)
seu <- FindClusters(seu, resolution = 0.3)
DimPlot(seu)
DimPlot(seu, group.by = "Line")
DimPlot(seu, group.by = "DiseaseStatus")

```


Annotate

gene lists 
```{r}

da_neurons <- c("TH","SLC6A3","SLC18A2","SOX6","NDNF","SNCG","ALDH1A1","CALB1","TACR2","SLC17A6","SLC32A1","OTX2","GRP","LPL","CCK","VIP")
NPC_orStemLike <- c("DCX","NEUROD1","TBR1","PCNA","MKI67","SOX2","NES","PAX6")
mature_neurons = c("RBFOX3","SYP","VAMP1","VAMP2","TUBB3","SYT1","BSN","HOMER1","SLC17A6")
excitatory_neurons = c("GRIA2","GRIA1","GRIA4","GRIN1","GRIN2B","GRIN2A","GRIN3A","GRIN3","GRIP1","CAMK2A")
inhbitory_neurons = c("GAD1","GAD2", "GAT1","PVALB","GABR2","GABR1","GBRR1","GABRB2","GABRB1","GABRB3","GABRA6","GABRA1","GABRA4","TRAK2")
astrocytes <- c("GFAP","S100B","AQP4","APOE", "SOX9","SLC1A3")
oligodendrocytes <- c("MBP","MOG","OLIG1","OLIG2","SOX10")

radial_glia <- c("PTPRC","AIF1","ADGRE1", "VIM", "TNC","PTPRZ1","FAM107A","HOPX","LIFR",
              "ITGB5","IL6ST","SLC1A3")
epithelial <- c("HES1","HES5","SOX2","SOX10","NES","CDH1","NOTCH1")

microglia <- c("IBA1","P2RY12","P2RY13","TREM119", "GPR34","SIGLECH","TREM2",
               "CX3CR1","FCRLS","OLFML3","HEXB","TGFBR1", "SALL1","MERTK",
               "PROS1")

features_list <- c("MKI67","SOX2","POU5F1","DLX2","PAX6","SOX9","HES1","NES","RBFOX3","MAP2","NCAM1","CD24","GRIA2","GRIN2B","GABBR1","GAD1","GAD2","GABRA1","GABRB2","TH","ALDH1A1","LMX1B","NR4A2","CORIN","CALB1","KCNJ6","CXCR4","ITGA6","SLC1A3","CD44","AQP4","S100B", "PDGFRA","OLIG2","MBP","CLDN11","VIM","VCAM1")

short_list <- c("MKI67","SOX9","HES1","NES","DLX2","RBFOX3","MAP2","TH","CALB1","KCNJ6","SLC1A3","CD44","AQP4","S100B","OLIG2","MBP","VIM")


gene_lists = list("DA_neurons" = da_neurons, "NPC" = NPC_orStemLike,
                  "Neurons" = mature_neurons, 
                  "Oligo" = oligodendrocytes, "RadialGlia" = radial_glia,
                  "Epithelial" = epithelial)
```

```{r}
Idents(seu) <- "integrated_snn_res.0.3"

for (i in da_neurons) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}
```

Make a dotplot of DA genes that are expressed

```{r}
da_neurons <- c("TH","SLC18A2","SOX6","NDNF","ALDH1A1","SLC17A6","SLC32A1","OTX2","LPL")

DotPlot(seu, features = da_neurons) + RotatedAxis()


```

```{r}
DefaultAssay(seu) <- "RNA"
seu <- ScaleData(seu)
Idents(seu) <- "integrated_snn_res.0.3"

for (i in NPC_orStemLike) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}
```
```{r}
Idents(seu) <- "integrated_snn_res.0.3"

for (i in mature_neurons) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}


```

```{r}
Idents(seu) <- "integrated_snn_res.0.3"

for (i in excitatory_neurons) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}


```

```{r}

Idents(seu) <- "integrated_snn_res.0.3"
inhbitory_neurons = c("GAD1","GAD2","PVALB","GABRB2","GABRB1","GABRB3","GABRA6","GABRA1","GABRA4","TRAK2")

for (i in inhbitory_neurons) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}

```

```{r}
Idents(seu) <- "integrated_snn_res.0.3"

for (i in epithelial) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}

```
```{r}
Idents(seu) <- "integrated_snn_res.0.3"
microglia <- c("P2RY12","P2RY13", "GPR34","TREM2",
               "CX3CR1","OLFML3","HEXB","TGFBR1", "SALL1","MERTK",
               "PROS1")
# no IBA1, TREM119"SIGLECH","FCRLS",
for (i in microglia) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}

```
```{r}
for (i in astrocytes) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}
```

```{r}
for (i in oligodendrocytes) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}
```
```{r}
for (i in radial_glia) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}
```

Get markers and check EnrichR

```{r}

ClusterMarkers <- FindAllMarkers(seu, only.pos = TRUE)

```

```{r}
library(enrichR)
setEnrichrSite("Enrichr") # Human genes
# list of all the databases
# get the possible libraries
dbs <- listEnrichrDbs()

# this will list the possible libraries
dbs

# select libraries with cell types
db <- c('CellMarker_Augmented_2021','Azimuth_Cell_Types_2021')

# function for a quick look
checkCelltypes <- function(cluster_num = 0){
  clusterX <- ClusterMarkers %>% filter(cluster == cluster_num & avg_log2FC > 0.25)
  genes <- clusterX$gene
  # the cell type libraries
  # get the results for each library
  clusterX.cell <- enrichr(genes, databases = db)
  # visualize the results
print(plotEnrich(clusterX.cell[[1]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value", title = 'CellMarker_Augmented_2021'))
print(plotEnrich(clusterX.cell[[2]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value", title = 'Azimuth_Cell_Types_2021'))

}

```
Top markers

```{r}
head(ClusterMarkers)
top5 <- ClusterMarkers %>% group_by(cluster) %>% top_n(n=5, wt =avg_log2FC)

DefaultAssay(seu) <- "RNA"
seu <- NormalizeData(seu)


top5

top2 <- ClusterMarkers %>% group_by(cluster) %>% top_n(n=2, wt =avg_log2FC)
DoHeatmap(seu, features = top5$gene, size = 3, angle = 90, group.by = "integrated_snn_res.0.3")

DoHeatmap(seu, features = top2$gene, size = 3, angle = 90, group.by = "integrated_snn_res.0.3")
```

quick check EnrichR

```{r}

checkCelltypes(cluster_num = 11)

```
Add annotations

```{r}

library(CelltypeR)
celltypes1 <- c("Neuron1","Neuron2","Epithelial","Neuron3", "Astro","Neuron4-TH","Neuron5",
                "Stem","Neuron6","Neuron7", "Pericyte", "Neuron8")

seu <- annotate(seu, annotations = celltypes1, to_label = "integrated_snn_res.0.3",annotation_name = "Celltypes1")
DimPlot(seu, label = TRUE, group.by = "Celltypes1")

celltypes2 <- c("imNeuron-DA","imNeuron-GABA-DA","Epithelial","imNeuron-DA",
                "Astro","imNeuron-Glut-DA","imNeuron-DA",
                "Stem","imNeuron-Glut","imNeuron-DA", "Pericyte", "NPC")

seu <- annotate(seu, annotations = celltypes2, to_label = "integrated_snn_res.0.3",annotation_name = "Celltypes2")
Idents(seu) <- "Celltypes2"
DimPlot(seu, label = TRUE)

celltypes3 <- c("imNeuron-DA","imNeuron-DA","Epithelial","imNeuron-DA",
                "Astro","imNeuron-DA","imNeuron-DA",
                "Stem","imNeuron","imNeuron-DA", "Pericyte", "NPC")

seu <- annotate(seu, annotations = celltypes3, to_label = "integrated_snn_res.0.3",annotation_name = "Celltypes3")
DimPlot(seu, label = TRUE)

```

```{r}
DimPlot(seu, label = FALSE, group.by = "DiseaseStatus")
```



Save the annotated neurons

```{r}

saveRDS(seu, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/Ex2NeuronsIntPDandHC.RDS")

```


