```{r}
library(Seurat)
library(dplyr)
library(Matrix)
library(ggplot2)

rm(list = ls())

install.packages("BiocManager")

BiocManager::install("clusterProfiler")
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("AnnotationDbi")
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("org.Hs.eg.db")

library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationDbi)
```
```{r}
SaveFigure <- function(plots, name, type = "png", width, height, res){
  if(type == "png") {
    png(paste0(fig_path, name, ".", type),
      width = width, height = height, units = "in", res = 200)
  } else {
    pdf(paste0(fig_path, name, ".", type),
      width = width, height = height)
}
print(plots)
dev.off()
}

SaveObject <- function(object, name){
  saveRDS(object, paste0(data_path, name, ".RDS"))
}

ReadObject <- function(name){
  readRDS(paste0(data_path, name, ".RDS"))
}
```

```{r}
mat_path <- "/Users/tgolds3/Documents/scRNA_seq/DGE_filtered"
mat <- ReadParseBio(mat_path)
```

```{r}
table(rownames(mat) == "")
rownames(mat)[rownames(mat) == ""] <- "unknown"
```

```{r}
cell_meta <- read.csv(paste0(mat_path, "/cell_metadata.csv"), row.names = 1)
```

```{r}
data48 <- CreateSeuratObject(mat, min_genes = 250, min_cells = 10,
names.feild = 0, meta.data = cell_meta)
```

```{r}
data48@meta.data$orig.ident <- factor(rep("data48", nrow(data48@meta.data)))
Idents(data48) <- data48@meta.data$orig.ident

SaveObject(data48, "seurat_obj_before_QC")
data48 <- ReadObject("seurat_obj_before_QC")
```
```{r}
data48[["percent.mt"]] <- PercentageFeatureSet(data48, pattern = "^MT-")
plot <- VlnPlot(data48, pt.size = 0.10, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
#SaveFigure(plot, "vln_QC", width = 12, height = 6)

plot1 <- FeatureScatter(data48, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(data48, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
# SaveFigure((plot1 + plot2),"scatter_QC", width = 12, height = 6, res = 200)
plot
plot1
plot2
```

```{r}
VlnPlot(data48, pt.size = 0.010, features = "nFeature_RNA", group.by = 'sample')
VlnPlot(data48, pt.size = 0.010, features = "nFeature_RNA", group.by = 'sample', y.max = 1000)

VlnPlot(data48, pt.size = 0.0010, features = "nCount_RNA", group.by = 'sample',y.max = 50000)
```

```{r}
data48 <- subset(data48, subset = nFeature_RNA < 12000 &  nFeature_RNA >300 & nCount_RNA < 30000 & percent.mt < 15)
```

```{r}
table(data48$sample)
```

```{r}
data48 <- NormalizeData(data48, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
data48 <- FindVariableFeatures(data48, selection.method = "vst", nfeatures = 2500)
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(data48), 10)
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(data48)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
#SaveFigure((plot1 + plot2), "var_features", width = 12, height = 6)
plot1
plot2
```

```{r}
data48 <- ScaleData(data48)
```

```{r}
data48 <- RunPCA(data48)

# Examine and visualize PCA results a few different ways
print(data48[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
plot <- VizDimLoadings(data48, dims = 1:2, reduction = "pca")
#SaveFigure(plot, "viz_PCA_loadings", width = 10, height = 8)
plot
```

```{r}
plot <- DimPlot(data48, reduction = "pca", group.by = "orig.ident")
#SaveFigure(plot, "pc1_2_scatter", width = 8, height = 6)
plot
```

```{r}
plot <- DimHeatmap(data48, dims = 1, cells = 500, balanced = TRUE, fast = FALSE)
#SaveFigure(plot, "dim_heatmap1", width = 8, height = 6)
plot
```

```{r}
plot <- DimHeatmap(data48, dims = 1:15, cells = 500, balanced = TRUE, fast = FALSE)
```
```{r}
plot <- ElbowPlot(data48,ndims = 50)
plot
```

```{r}
data48 <- FindNeighbors(data48, dims = 1:30)
data48 <- FindClusters(data48, resolution = 0.30)
```

```{r}
data48 <- BuildClusterTree(data48, reorder = TRUE, reorder.numeric = TRUE)
```

```{r}
data48 <- RunUMAP(data48, dims = 1:30)
plot <- DimPlot(data48, reduction = "umap")
plot
```

```{r}
DimPlot(data48, reduction = "umap", group.by = "sample")
```

```{r}
mar <- FindAllMarkers(data48, min.pct = 0.25, logfc.threshold = 0.25)
mar %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC)

mar
```


```{r}
write.csv(mar, "/Users/tgolds3/Documents/scRNA_seq/MarkersforClusters.csv")
```

```{r}
top5 <- mar %>% group_by(cluster)
to_plot <- unique(top5$gene)
plot <- DotPlot(data48, features = to_plot, group.by = "tree.ident")
#SaveFigure(plot, "dplot_top5", width = 9, height = 20)
plot
```

```{r}
saveRDS(data48,"/Users/tgolds3/Documents/scRNA_seq/Parse48_analysis.RDS")
```
```{r}
genes_to_test <- rownames(mar[mar$log2FoldChange > 0.5])
genes_to_test
```
```{r}
GO_results <- enrichGO(gene = genes_to_test, OrgDb = "org.Hs.eg.db", keyType = "SYMBOL", ont = "BP")
as.data.frame(GO_results)
```

```{r}
write.csv(GO_results, "/Users/tgolds3/Documents/scRNA_seq/GO_Biological_Processes.csv", row.names=FALSE)
```

```{r}
fit <- plot(barplot(GO_results, showCategory = 10))
```
Elia said this is how you can subset data! 
sdata <- subset(data48, sample %in% c("M1", "M2", "M3", "M4", "M5", "M6", "M7", "M8"))
This was run in the terminal, so if I want to do it in the block need to run it alone in a code block thing. 

Now Running all steps above for analysis on maria's samples
```{r}
sdata <- FindNeighbors(sdata, dims = 1:30)
sdata <- FindClusters(sdata, resolution = 0.30)

sdata <- BuildClusterTree(sdata, reorder = TRUE, reorder.numeric = TRUE)
```


```{r}
sdata <- RunUMAP(sdata, dims = 1:30)
plot_maria_umap <- DimPlot(sdata, reduction = "umap")
plot_maria_umap

DimPlot(sdata, reduction = "umap", group.by = "sample")

mm <- FindAllMarkers(sdata, min.pct = 0.25, logfc.threshold = 0.25)

mm %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC)

mm

write.csv(mm, "/Users/tgolds3/Documents/scRNA_seq/Maria_MarkersforClusters.csv")

top5m <- mm %>% group_by(cluster)
to_plotm <- unique(top5m$gene)
plot_m <- DotPlot(sdata, features = to_plotm, group.by = "tree.ident")
#SaveFigure(plot, "dplot_top5", width = 9, height = 20)
plot_m

genes_to_test_maria <- rownames(mm[mm$log2FoldChange > 0.5])
genes_to_test_maria

GO_results_maria <- enrichGO(gene = genes_to_test_maria, OrgDb = "org.Hs.eg.db", keyType = "SYMBOL", ont = "BP")
as.data.frame(GO_results_maria)

write.csv(GO_results_maria, "/Users/tgolds3/Documents/scRNA_seq/Maria_GO_Biological_Processes.csv", row.names=FALSE)

fitm <- plot(barplot(GO_results_maria, showCategory = 10))





```

```{r}
function(fitm, MariaGO, type = "png", width, height, res){
    if(type == "png") {
        png(paste0("Users/tgolds3/Documents/scRNA_seq", MariaGO, ".", type),
            width = 5, height = 8, units = "in", res = 300)
    } else {
        pdf(paste0(fig_path, name, ".", type),
            width = width, height = height)
    }
    print(fitm)
    dev.off()
}
```
# print GO

pdf("MariaGo.pdf")
print(fitm)
dev.off()


Got some more graphs for Maria and fixed an issue
DimPlot(sdata, reduction = "umap")
> plot_maria_umap
> plot_maria_umap <- DimPlot(sdata, reduction = "umap")
> plot_maria_umap
> pdf("Maria_umap.pdf")
> print(plot_maria_umap)
> dev.off()
RStudioGD 
        2 
> DimPlot(sdata, reduction = "umap", group.by = "sample")
> sampleumapmaria <- DimPlot(sdata, reduction = "umap", group.by = "sample")
> pdf("UMAP_Sample_Maria.pdf")
> print(sampleumapmaria)
> dev.off()
RStudioGD 
        2 
> pdf("All_UMAP_Maria.pdf")
> print(plot_maria_umap)
> print(sampleumapmaria)
> dev.off()
RStudioGD 
        2 
> View(sdata)