---
title: "R Notebook"
output: html_notebook
---

Prepare data for NCADD sharing

Data objects 
• Experiment 1: 12 samples NPCs (4 lines X 3 replicates)
• Experiment 2: NPCs (4 HC, 8 PD, no replicates)
• Experiment 1 + 2 : 12 samples NPCs (4 lines X 3 replicates)NPCs (4 HC, 8 PD, no replicates)
* Experiment 2: Neurons (2 HC, 4 PD, no replicates)

```{r}
library(Seurat)
library(tidyverse)

```

The same annotation should be used on the two NPC data types

```{r}
NPCex1 <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Experiment1-mini12/Integrated12samples.RDS")

NPCex2 <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesintegratedNPCs.RDS")

colnames(NPCex1@meta.data)
colnames(NPCex2@meta.data)

# remove the isogenics
#Idents(NPCex2) <- "DiseaseStatus"
#seu <- subset(NPCex2, idents = c("HC","PD"))
#unique(seu$DiseaseStatus)

#NPCex2 <- seu
#rm(seu)
#unique(NPCex2$DiseaseStatus)

#unique(NPCex1$Celltypes1)
#unique(NPCex2$Celltypes1)

#saveRDS(NPCex2, "/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesintegratedNPCsPDHConly.RDS")

# NPCex2 <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesintegratedNPCsPDHConly.RDS")

```

Integrate the 2 NPC samples

```{r}

dim(NPCex1)
dim(NPCex2)

```

Integrate

```{r}

DefaultAssay(NPCex1) <- "RNA"
DefaultAssay(NPCex2) <- "RNA"
NPCex1$orig.ident <- "NPCex1"
NPCex2$orig.ident <- "NPCex2"

sublist <- list("EX1" = NPCex1,"EX2" = NPCex2)
seu <- integrate_seurat_objects(sublist)


```

Try to break apart the steps
```{r}

seurat_list <- sublist
rm(sublist)

for (i in 1:length(seurat_list)) {
    seurat_list[[i]] <- NormalizeData(seurat_list[[i]], verbose = FALSE)
    seurat_list[[i]] <- ScaleData(seurat_list[[i]], verbose = FALSE)
    seurat_list[[i]] <- FindVariableFeatures(seurat_list[[i]], selection.method = "vst")
    seurat_list[[i]] <- RunPCA(seurat_list[[i]], verbose = FALSE, npcs = 30)
  }
  # Find integration anchors
  int.anchors <- FindIntegrationAnchors(object.list = seurat_list, dims = 1:30,
                                        anchor.features = 1000)

#saveRDS(int.anchors,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/AnchorsNPCex1and2.RDS")
  
# read in the anchors 
int.anchors <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/AnchorsNPCex1and2.RDS")

#saveRDS(seurat_list,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ListNPCex1and2.RDS")
  # Integrate data
seurat_list <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ListNPCex1and2.RDS")

seu <- merge(seurat_list[[1]], seurat_list[[2]])
seu <- IntegrateData(anchorset = int.anchors)



  
  

```




```{r}
# make UMAP from integrated
seu <- integrated_seurat
seu <- ScaleData(seu)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:20, n.neighbors = 81)
seu <- FindVariableFeatures(seu)
seu <- FindNeighbors(seu, dims = 1:20, k.param = 81)
seu <- FindClusters(seu, resolution = 0.3)
DimPlot(seu)

```

Look at the current cell type annotations 

```{r}

DimPlot(seu, group.by = "Celltypes1")

```



```{r}
tb <- table(NPCex2$DiseaseStatus,NPCex2$Line)
print(tb)
length(unique(NPCex2$Line))
input_vector <- unique(NPCex2$Line)

unique(NPCex2$sample)



```
```{r}
neurons <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesNeurons.RDS")

```


```{r}
unique(neurons$sample)

```


Test removing excess metadata

```{r}
# Assuming you have a Seurat object named 'seurat_obj'
# And you want to remove metadata columns except for the ones listed in 'metadata_to_keep'
seurat_obj <- NPCex2
# Define the metadata columns you want to keep
metadata_to_keep <- c("orig.ident","nCount_RNA","nFeature_RNA","gene_count",
                      "sample","DiseaseStatus", "Line","Celltypes1",
                      "Celltypes2","Celltypes3")

# Update the Seurat object to keep only the desired metadata columns
seurat_obj@meta.data <- seurat_obj@meta.data[, metadata_to_keep]
colnames(seurat_obj@meta.data)

# add the Batch annotation
Idents(seurat_obj) <- "sample"
levels(seurat_obj)



```
```{r}
annotate <- function(seu, annotations, to_label, annotation_name = "CellType"){
  Idents(seu) <- to_label
  names(annotations) <- levels(seu)
  seu <- RenameIdents(seu, annotations)
  seu <- AddMetaData(object=seu, metadata=Idents(seu), col.name = annotation_name)

}

```

```{r}
# add batch
# batch 4 for replicate 1 and batch 5 if there is a second replicate

Batch <- c("B4", "B4",  "B5", "B4", "B4", "B4",
           "B4", "B4",  "B4",  "B4",   "B4", "B4",
           "B4", "B4",  "B4",  "B4", "B4",  "B4",
           "B4",  "B4",  "B4",   "B4", "B4")

seurat_obj <- annotate(seurat_obj, annotations = Batch, to_label = "sample",
                       annotation_name = "Batch")

levels(seurat_obj)

DimPlot(seurat_obj, group.by = "sample")
DimPlot(seurat_obj, group.by = "DiseaseStatus")

Idents(seurat_obj) <- "DiseaseStatus"

NPCex2 <- subset(seurat_obj, idents = c("HC","PD"))

DimPlot(NPCex2, group.by = "DiseaseStatus")

```


```{r}
# save the EX2 NPC with only the PD and HC plus the batch annotation
saveRDS(NPCex2,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/Ex2_48well/ParseNCADDsamplesintegratedNPCsPDHConly.RDS")

```

Check the orig.ident
```{r}

Idents(NPCex2) <- "orig.ident"
NPCex2$orig.ident <- "EX2"
unique(NPCex2$orig.ident)

```

Look at the EX1 and remove excess meta data

```{r}
colnames(NPCex1@meta.data)
NPCex1 <- AddMetaData(NPCex1, metadata = Idents(NPCex1), col.name = "Line")

seurat_obj <- NPCex1
# Define the metadata columns you want to keep
metadata_to_keep <- c("orig.ident","nCount_RNA","nFeature_RNA","gene_count",
                      "sample","DiseaseStatus", "Line","Celltypes1",
                      "Celltypes2","Celltypes3")

# Update the Seurat object to keep only the desired metadata columns
seurat_obj@meta.data <- seurat_obj@meta.data[, metadata_to_keep]
colnames(seurat_obj@meta.data)

Idents(seurat_obj) <- "orig.ident"
seurat_obj$orig.ident <- "EX1"
unique(seurat_obj$orig.ident)

DimPlot(seurat_obj, group.by = "DiseaseStatus")

```

Process the two data objects

```{r}
# EX1
NPCex1 <- seurat_obj

NPCex1 <- NormalizeData(NPCex1, verbose = FALSE)
NPCex1 <- ScaleData(NPCex1, verbose = FALSE) # must be run for PCA?
NPCex1 <- FindVariableFeatures(NPCex1, selection.method = "vst")
NPCex1 <- RunPCA(NPCex1, verbose = FALSE, npcs = 30)
NPCex1 <- RunUMAP(NPCex1, dims = 1:30)

DimPlot(NPCex1, group.by = "Celltypes1")
DimPlot(NPCex1, group.by = "Celltypes2")
DimPlot(NPCex1, group.by = "Celltypes3")

```


```{r}

NPCex2 <- NormalizeData(NPCex2, verbose = FALSE)
NPCex2 <- ScaleData(NPCex2, verbose = FALSE)
NPCex2 <- FindVariableFeatures(NPCex2, selection.method = "vst")
NPCex2 <- RunPCA(NPCex2, verbose = FALSE, npcs = 30)
NPCex2 <- RunUMAP(NPCex2, dims = 1:30)

DimPlot(NPCex2, group.by = "Celltypes1")
DimPlot(NPCex2, group.by = "Celltypes2")
DimPlot(NPCex2, group.by = "Celltypes3")

```

```{r}
NPCex1 <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/NPCex1.RDS")
NPCex2 <- readRDS("/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/NPCex2.RDS")
```


Integrate the seurat objects

```{r}

seu <- merge(NPCex1, NPCex2)
unique(seu$orig.ident)

```


Save the processed objects

```{r}
#saveRDS(NPCex1,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/NPCex1.RDS")
#saveRDS(NPCex2,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/NPCex2.RDS")

saveRDS(seu,"/Users/rhalenathomas/Documents/Data/scRNAseq/ParseExample/NPCex1_ex2.RDS")


```



See a UMAP without integration just from the merged but using the integrated RNA from the previous integrations


```{r}
DefaultAssay(seu) <- "integrated"

seu <- ScaleData(seu, verbose = FALSE)
int_genes <- rownames(seu@assays[["integrated"]]@meta.features)
#seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000)
# use the genes that were for integrating each of the separate data sets
# each would have been set to 2000 and there is 2427 genes - many will be overlapping
seu <- RunPCA(seu, npcs = 30, verbose = FALSE, features = int_genes)
seu <- RunUMAP(seu, reduction = "pca", dims = 1:30)

DimPlot(seu, group.by = "orig.ident")

# okay so they need to be integrated!

```
```{r}

# the variable features were removed either by merging or by reprocessing for PCA 
seu <- merge(NPCex1, NPCex2)

seurat_list <- SplitObject(seu, split.by = "orig.ident")

# Find integration anchors
int.anchors <- FindIntegrationAnchors(object.list = seurat_list, dims = 1:25,
                                        anchor.features = 1000)

# Integrate the data
integrated_seurat <- IntegrateData(anchorset = integration_anchors, dims = 1:25)


```


